Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data

### 2021-07-25 Manuscript/00-main.Rmd
started flow chart
Added Flowchart of Prior

---
title: Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data
author:
- name: Boyi Guo*
  num: a
- name: Byron C. Jaeger
  num: b 
- name: AKM Fazlur Rahman
  num: a   
- name: D. Leann Long
  num: a  
- name: Nengjun Yi
  num: a
address:
- num: a
  org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA
- num: b
  org: Department of Biostatistics and Data Science, Wake Forest University, Winston-Salem, USA
corres: "*Corresponding author name, This is sample corresponding address. \\email{xxx@uab}"
presentaddress: This is sample for present address text this is sample for present address text
authormark: Guo et al.
articletype: Research article
received: 2021-01-01
revised: 2021-02-01
accepted: 2021-03-01
abstract: "Spike-and-slab priors have been extensively studied for variable selection, and recently been extended to function selection in generalized additive models (GAMs). Most of the previous works on fitting spike-and-slab GAM relies on Markov chain Monte Carlo (MCMC) algorithms to identify the promising subsets of functions. The computational burden of MCMC makes these models less feasible, especially in high-dimensional setting. Besides, most of the current spike-and-slab GAM consider an “all-in-all-out” approach for function selection, rendering less flexibility to answer if non-linear effects are necessary in the model. We propose Bayesian hierarchical generalized additive models for function selection and prediction. The proposed models employ spike-and-slab priors, mixture normal and mixture double exponential, for smooth and sparse solution. Two EM based algorithms, EM-IWLS and EM-Coordinate Descent, are developed for accelerated computation. Simulation studies and real data analysis demonstrated improved performance against the state-of-art models, mgcv and COSSO. The software implementation of the proposed methods is freely available via the R package BHAM."
keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent
bibliography: bibfile.bib
output: 
  rticles::sim_article:
    keep_tex: TRUE
---

\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\tp}{\text{pen}}

>>>>

---
title: Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data
author:
- name: Boyi Guo*
  num: a
- name: Byron C. Jaeger
  num: b 
- name: AKM Fazlur Rahman
  num: a   
- name: D. Leann Long
  num: a  
- name: Nengjun Yi
  num: a
address:
- num: a
  org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA
- num: b
  org: Department of Biostatistics and Data Science, Wake Forest University, Winston-Salem, USA
corres: "*Corresponding author name, This is sample corresponding address. \\email{xxx@uab}"
presentaddress: This is sample for present address text this is sample for present address text
authormark: Guo et al.
articletype: Research article
received: 2021-01-01
revised: 2021-02-01
accepted: 2021-03-01
abstract: "Spike-and-slab priors have been extensively studied for variable selection, and recently been extended to function selection in generalized additive models (GAMs). Most of the previous works on fitting spike-and-slab GAM relies on Markov chain Monte Carlo (MCMC) algorithms to identify the promising subsets of functions. The computational burden of MCMC makes these models less feasible, especially in high-dimensional setting. Besides, most of the current spike-and-slab GAM consider an “all-in-all-out” approach for function selection, rendering less flexibility to answer if non-linear effects are necessary in the model. We propose Bayesian hierarchical generalized additive models for function selection and prediction. The proposed models employ spike-and-slab priors, mixture normal and mixture double exponential, for smooth and sparse solution. Two EM based algorithms, EM-IWLS and EM-Coordinate Descent, are developed for accelerated computation. Simulation studies and real data analysis demonstrated improved performance against the state-of-art models, mgcv and COSSO. The software implementation of the proposed methods is freely available via the R package BHAM."
keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent
bibliography: bibfile.bib
output: 
  rticles::sim_article:
    keep_tex: TRUE
header-includes: 
  - \usepackage{tikz}
  - \usepackage{pgfplots}
---

<!-- \usepackage{tikz} -->
\usetikzlibrary{shapes.geometric, arrows, positioning, calc, matrix, backgrounds, fit}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\tp}{\text{pen}}

### 2021-07-25 Manuscript/02-Method.Rmd
Revise the simulation section
Added Flowchart of Prior

This is to leverage the fact that the probability of selecting the bases of a smooth function should be similar, while allowing different penalty on the null space and penalty space of the spline function. The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level. Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^\tp$ are independently distributed for analytic simplicity. We further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
$$
\theta_j \sim Beta(a, b).
$$
The beta distribution is a conjugate prior for the binomial distribution and hence provides some computation convenience. For simplicity, we focus on a special case of beta distribution, uniform (0,1). When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases. Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL.

```{r fig.height=4, fig.width=6}
# TODO: Refactor the graph to a latex flowchart
include_graphics("Fig/SS_prior_flow_chart.png")
```

@@

Alternative to the sampling algorithms for fitting Bayesian models, optimization algorithms focus on the maximum a posteriori (MAP) estimates and speed up the process at the cost of uncertainty measures. The earliest work for fitting spike-and-slab models using optimization algorithm is EMVS [@Rockova2014a]. @Rockova2014a proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors. 
<!-- The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$. The parameter $\sigma^2$ follows a inverse gamma prior, $IG(v/2, v\lambda /2)$. Like in the spike-and slab family, the latent binary variables $\bs \gamma$ follows a Bernoulli prior with a hyper prior $\theta$ follows a beta distribution, or simpler the uniform (0,1).  -->
In the E step, the latent variables $\gamma$ are treated as the missing data, and calculate the posterior mean of the density function; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly. 
<!-- As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter. EMVS is also compatible with structured prior when dealing with grouped variables. The group structure can be imposed on the prior of the latent binary variable $\gamma$ via either logistic regression or markov random field.  -->
The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coordinate descent (CD) algorithm for fitting SSL models.
<!-- the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL). The authors also lay the theoretical discussion of comparing spike-and-slab from the penalty perspective and the Bayesian perspective. The SSL also mitigated the problem of EMVS that the weak signals are not shrink to zero. -->
Yi and his group incorporated the SSL prior in the EM-Iterative Weighted Least Square(IWLS) algorithm and proposed another EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty SSL exhibit. Such applications were used in generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019]. Both EM based algorithms provide deterministic solutions, which becomes a more popular property as reproducible results. Bai et al. [@Bai2020] provides an up-to-date summary of SSL methods and theoretical discussion.

>>>>

This is to leverage the fact that the probability of selecting the bases of a smooth function should be similar, while allowing different penalty on the null space and penalty space of the spline function. The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level. Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^\tp$ are independently distributed for analytic simplicity. We further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
$$
\theta_j \sim Beta(a, b).
$$
The beta distribution is a conjugate prior for the binomial distribution and hence provides some computation convenience. For simplicity, we focus on a special case of beta distribution, uniform (0,1). When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases. Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL. \\

```{r fig.height=4, fig.width=6, eval = F}
# TODO: Refactor the graph to a latex flowchart
include_graphics("Fig/SS_prior_flow_chart.png")
```

\begin{tikzpicture} [
staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
]

\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
&\node (linearTheta) [outCome] { $\theta_j^{0}\sim Beta(a, b)$ };\\
&\node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j^0) $ };\\
&\node (linearBeta) [outCome] { $\beta_j^0 \sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
};

\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
\node (penTheta) [outCome] { $\theta_{j}^\tp\sim Beta(a, b)$ };\\
\node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \theta_j^\tp)$ };\\
\node (penBeta) [outCome] { $\beta_{jk}^\tp \sim  DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
};

\node (ab)[staticCompo] at ($(linearTheta)!0.5!(penTheta)$) {$(a, b)$};
%\node (PenaltyPrior) [PriorBoarder,  right = 1cm of ab] {part 2};
\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};

\node (Beta) [staticCompo, below = 1cm of s] {$\bs \beta = (\beta^0_1, \bs \beta^\tp_1, \dots,\beta^0_j, \bs \beta^\tp_j , \dots,\beta^0_p, \bs \beta^\tp_p) $};
\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g(\eta_i))$};

%Lines

\draw[->] (linearTheta) -- (linearGamma);
\draw[->] (penTheta) -- (penGamma);
\draw[->] (linearGamma) -- (linearBeta) ;
\draw[->] (penGamma) -- (penBeta);
\draw[->] (ab) -- (linearTheta);
\draw[->] (ab) -- (penTheta);
\draw[->] (s) -- (linearBeta) ;
\draw[->] (s) -- (penBeta);
\draw[->] (linearBeta) -- (Beta);
\draw[->] (penBeta) -- (Beta);
\draw[->] (Beta) --  (Y);

% Now that the diagram has been drawn, background rectangles
% can be fitted to its elements. This requires the TikZ
% libraries "fit" and "background".
% Control input and measurement are labeled. These labels have
% not been translated to English as "Measurement" instead of
% "Messung" would not look good due to it being too long a word.
\begin{pgfonlayer}{background}
    \node [background,
                fit=(linearTheta) (linearBeta),
                label=above:Null Space:] {};
    \node [background,
                 fit=(penTheta) (penBeta),
                label=above:Penalized Space:] {};

\end{pgfonlayer}
\end{tikzpicture}

@@

Alternative to the sampling algorithms for fitting Bayesian models, optimization algorithms focus on the maximum a posteriori (MAP) estimates and speed up the process at the cost of uncertainty measures. The earliest work for fitting spike-and-slab models using optimization algorithm is EMVS [@Rockova2014a]. Ro{\v{c}}kov{\'{a}} and George[@Rockova2014a] proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors. 
<!-- The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$. The parameter $\sigma^2$ follows a inverse gamma prior, $IG(v/2, v\lambda /2)$. Like in the spike-and slab family, the latent binary variables $\bs \gamma$ follows a Bernoulli prior with a hyper prior $\theta$ follows a beta distribution, or simpler the uniform (0,1).  -->
In the E step, the latent variables $\gamma$ are treated as the missing data, and calculate the posterior mean of the density function; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly. 
<!-- As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter. EMVS is also compatible with structured prior when dealing with grouped variables. The group structure can be imposed on the prior of the latent binary variable $\gamma$ via either logistic regression or markov random field.  -->
The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coordinate descent (CD) algorithm for fitting SSL models.
<!-- the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL). The authors also lay the theoretical discussion of comparing spike-and-slab from the penalty perspective and the Bayesian perspective. The SSL also mitigated the problem of EMVS that the weak signals are not shrink to zero. -->
Yi and his group incorporated the SSL prior in the EM-Iterative Weighted Least Square(IWLS) algorithm and proposed another EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty SSL exhibit. Such applications were used in generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019]. Both EM based algorithms provide deterministic solutions, which becomes a more popular property as reproducible results. Bai et al. [@Bai2020] provides an up-to-date summary of SSL methods and theoretical discussion.

### 2021-07-25 Manuscript/SS_GAM.tex
Added Flowchart of Prior

@@

\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\tp}{\text{pen}}

@@

This is to leverage the fact that the probability of selecting the bases
of a smooth function should be similar, while allowing different penalty
on the null space and penalty space of the spline function. The hyper
prior of \(\gamma_{j}^{0}\) decides the sparsity of the model at the
function selection level, while that of \(\gamma_{j}^\text{pen}\)
decides the smoothness of the spline function at basis function level.
Meanwhile, we specify that \(\gamma_{j}^0\) and
\(\gamma_{j}^\text{pen}\) are independently distributed for analytic
simplicity. We further specify the parameter \(\theta_j\) follows a beta
distribution with given shape parameters \(a\) and \(b\), \[
\theta_j \sim Beta(a, b).
\] The beta distribution is a conjugate prior for the binomial
distribution and hence provides some computation convenience. For
simplicity, we focus on a special case of beta distribution, uniform
(0,1). When the variable have large effects in any of the bases, the
parameter \(\theta_j\) will be estimated large, which in turn encourages
the model to include the rest of bases, achieving the local adaption
among spline bases. Hereafter, we refer Bayesian hierarchical
generalized additive models with the spike-and-slab lasso prior as the
BHAM-SSL.

\includegraphics{Fig/SS_prior_flow_chart.png}

@@

In this section, we compare the performance of the proposed models to
four alternative models: component selection and smoothing operator
(COSSO) \citep{Zhang2006GAM}, adaptive COSSO \citep{Storlie2011},
generalized additive models with automatic smoothing \citep{Wood2011},
SB-GAM \citep{Bai2021}. COSSO is one of the earliest smoothing spline
models that consider sparsity-smoothness penalty. Adaptive COSSO
improved upon COSSO by using adaptive weight for penalties such that the
penalty of each functional component are different for extra
flexibility. Generalized additive models with automatic smoothing,
hereafter \textit{mgcv}, is one of the most popular models for
non-linear effect interpolation and prediction. SB-GAM is the first
spike-and-slab lasso GAM. We implemented COSSO and adaptive COSSO with R
package \texttt{cosso} \citep{cosso} 2.1-1, generalized additive models
with automatic smoothing with R package \texttt{mgcv} 1.8-31, SB-GAM
with R package \text{sparseGAM} 1.0. All computation was conducted on a
high-performance 64-bit Linux platform with 48 cores of 2.70GHz
eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R
3.6.2 \citep{R}.

@@

xxx method. Adaptive COSSO is another method, mgcv is another method. To
demonstrate the correctness and feasibility of the proposed SS-GAM, we
follow the logistic regression simulation introduced in Bai (Work in
Progress). We first generate \(n=500\) training data sample with number
of predictors \(p=4, 10, 50, 200\) respectively, where the predictors
\(X \sim MVN_{n\times p}(0, I_{P})\). We simulate the outcome data from
three distributions, gaussian, poisson, and binomial with the identity
link, logit link \(g(x) = \log(\frac{x}{1-x})\), and log link
\(g(x) = log(x)\) functions to mimic the common use of generalized
model. The outcome is simulated via the following function \[
\mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
\] for gaussian and binomial outcome and shrinked effects

\[
\mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3))
\] for stable poisson outcome. For the Gaussian outcome where dispersion
are required, the dispersion parameter was set to 1. We follow the same
data generating functions in Bai. In this data generating process, we
have \(x_1, x_2, x_3, x_4\) as the active covariates, while the rest
covariates are inactive \(f_j(x_j) = 0\) for \(j = 4, \dots, p\).
Another set of independent sample of size
\(n_{test}=1000\),(\(x_{new} ,y_{new}\)), are created following the same
data generating process, serving as the testing data. For mgcv and
proposed models where choice of spline function are customizable, we
choose to cubic regression spline with 10 basis. In COSSO and adaptive
COSSO where the choice of spline are pre-determined, we set to have 10
bases. The corresponding number of parameter that we estimate with the
model is 37, 82, 451 and 1801. We also run the simulation where the
number of bases are increased to 25 deliberately. To evaluate the
performance of the models, mean squared error, mean absolute error for
gaussian model, AUC and misclassifcation rate are calculated for
binomial model, and deviance are calculated for Poisson model, averaging
across 50 simulations. The results are tabulated in Table 1, 2, 3.

>>>>

\usepackage{tikz}
\usepackage{pgfplots}

@@

\usetikzlibrary{shapes.geometric, arrows, positioning, calc, matrix, backgrounds, fit}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\tp}{\text{pen}}

@@

This is to leverage the fact that the probability of selecting the bases
of a smooth function should be similar, while allowing different penalty
on the null space and penalty space of the spline function. The hyper
prior of \(\gamma_{j}^{0}\) decides the sparsity of the model at the
function selection level, while that of \(\gamma_{j}^\text{pen}\)
decides the smoothness of the spline function at basis function level.
Meanwhile, we specify that \(\gamma_{j}^0\) and
\(\gamma_{j}^\text{pen}\) are independently distributed for analytic
simplicity. We further specify the parameter \(\theta_j\) follows a beta
distribution with given shape parameters \(a\) and \(b\), \[
\theta_j \sim Beta(a, b).
\] The beta distribution is a conjugate prior for the binomial
distribution and hence provides some computation convenience. For
simplicity, we focus on a special case of beta distribution, uniform
(0,1). When the variable have large effects in any of the bases, the
parameter \(\theta_j\) will be estimated large, which in turn encourages
the model to include the rest of bases, achieving the local adaption
among spline bases. Hereafter, we refer Bayesian hierarchical
generalized additive models with the spike-and-slab lasso prior as the
BHAM-SSL. \textbackslash{}

\begin{tikzpicture} [
staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
]

\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
&\node (linearTheta) [outCome] { $\theta_j^{0}\sim Beta(a, b)$ };\\
&\node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j^0) $ };\\
&\node (linearBeta) [outCome] { $\beta_j^0 \sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
};

\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
\node (penTheta) [outCome] { $\theta_{j}^\text{pen}\sim Beta(a, b)$ };\\
\node (penGamma) [outCome] { $\gamma_{j}^\text{pen}\sim Bin(1, \theta_j^\text{pen})$ };\\
\node (penBeta) [outCome] { $\beta_{jk}^\text{pen}\sim  DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
};

\node (ab)[staticCompo] at ($(linearTheta)!0.5!(penTheta)$) {$(a, b)$};
%\node (PenaltyPrior) [PriorBoarder,  right = 1cm of ab] {part 2};
\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};

\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta^0_1, \boldsymbol{\beta}^\text{pen}_1, \dots,\beta^0_j, \boldsymbol{\beta}^\text{pen}_j , \dots,\beta^0_p, \boldsymbol{\beta}^\text{pen}_p) $};
\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g(\eta_i))$};

%Lines

\draw[->] (linearTheta) -- (linearGamma);
\draw[->] (penTheta) -- (penGamma);
\draw[->] (linearGamma) -- (linearBeta) ;
\draw[->] (penGamma) -- (penBeta);
\draw[->] (ab) -- (linearTheta);
\draw[->] (ab) -- (penTheta);
\draw[->] (s) -- (linearBeta) ;
\draw[->] (s) -- (penBeta);
\draw[->] (linearBeta) -- (Beta);
\draw[->] (penBeta) -- (Beta);
\draw[->] (Beta) --  (Y);

% Now that the diagram has been drawn, background rectangles
% can be fitted to its elements. This requires the TikZ
% libraries "fit" and "background".
% Control input and measurement are labeled. These labels have
% not been translated to English as "Measurement" instead of
% "Messung" would not look good due to it being too long a word.
\begin{pgfonlayer}{background}
    \node [background,
                fit=(linearTheta) (linearBeta),
                label=above:Null Space:] {};
    \node [background,
                 fit=(penTheta) (penBeta),
                label=above:Penalized Space:] {};

\end{pgfonlayer}
\end{tikzpicture}

@@

In this section, we compare the performance of the proposed models to
four alternative models: component selection and smoothing operator
(COSSO) \citep{Zhang2006GAM}, adaptive COSSO \citep{Storlie2011},
generalized additive models with automatic smoothing \citep{Wood2011},
SB-GAM \citep{Bai2021}. COSSO is one of the earliest smoothing spline
models that consider sparsity-smoothness penalty. Adaptive COSSO
improved upon COSSO by using adaptive weight for penalties such that the
penalty of each functional component are different for extra
flexibility. Generalized additive models with automatic smoothing,
hereafter \textit{mgcv}, is one of the most popular models for
non-linear effect interpolation and prediction. SB-GAM is the first
spike-and-slab lasso GAM. We implemented COSSO and adaptive COSSO with R
package \texttt{cosso 2.1-1}, generalized additive models with automatic
smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R package
\text{sparseGAM 1.0}. COSSO models and SB-GAM do not provide flexibility
to define smoothing functions, and hence used the default choices. Both
mgcv and proposed models allow customized smoothing functions and we
chose the cubic regression spline. We controlled the dimensionality of
each smoothing function, 10 bases, for all different choices of
smoothing functions. 5-fold cross-validation were used for COSSO models,
SB-GAM and the proposed models for tuning parameter selection based on
the default selection criteria. 20 default candidates were examined for
SB-GAM and the proposed models which allow user-specification of tuning
candidates. All computation was conducted on a high-performance 64-bit
Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680
processors and 24G of RAM per core and R 3.6.2 \citep{R}.

@@

We follow the data generating process described in Bai \citep{Bai2021}.
We first generated \(n=500\) training data points with
\(p=4, 10, 50, 200\) predictors respectively, where the predictors \(X\)
are simulated from a multivariate normal distribution
\(\text{MVN}_{n\times p}(0, I_{P})\). We then simulated the outcome
\(y\) from three distribution distributions, gaussian, binomial, and
poisson with the identity link, logit link
\(g(x) = \log(\frac{x}{1-x})\), and log link \(g(x) = log(x)\). The mean
of each outcome were simulated via the following function \[
\mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
\] for gaussian and binomial outcomes and shrinked effects

\[
\mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3))
\] for poisson outcomes. Gaussian outcomes required specification of
dispersion, where we set the dispersion parameter to be 1. In this data
generating process, we have \(x_1, x_2, x_3, x_4\) as the active
covariates, while the rest covariates are inactive,
i.e.~\(f_j(x_j) = 0\) for \(j = 4, \dots, p\). Another set of
independent sample of size \(n_{test}=1000\),(\(x_{new} ,y_{new}\)), are
created following the same data generating process, serving as the
testing data. The dimensionality of the smooth functions are controlled
as k=10. Hence, the corresponding number of coefficient parameters to be
estimated is 37, 82, 451 and 1801.

To evaluate the performance of the models, mean squared error, mean
absolute error for gaussian model, AUC and misclassifcation rate are
calculated for binomial model, and deviance are calculated for Poisson
model, averaging across 50 simulations. The results are tabulated in
Table 1, 2, 3.
